Arrows for Parallel Computation

### 2018-01-06 paper/content/motivation.tex
phil remarks pages 1-22 (ish)
minor

We provide an Arrows-based type class and implementations for the three above mentioned parallel Haskells.
Instead of 
introducing a new low-level parallel backend to implement our
Arrows-based interface, we define a shallow-embedded DSL for Arrows. This DSL
is defined as a common interface with varying implementations in
the existing parallel Haskells. Thus, we not only define a parallel programming interface in a
novel manner -- we tame the zoo of parallel Haskells. We provide a
common, very low-penalty programming interface that allows to switch
the parallel backends at will. The induced penalty was in the single-digit percent range, with means over the varying cores configuration typically under 1\% overhead in our measurements (Section~\ref{sec:benchmarks}). Further backends based on HdpH or a Frege implementation (on the Java Virtual Machine) are viable, too.

@@

We structure this paper as follows:
\begin{itemize}
\item We introduce a parallel evaluation formalism using Arrows. One big advantage of our specific approach is that we do not have to introduce any new types, facilitating composability (Section~\ref{sec:parallel-arrows}).
\item We show that PArrow programs can readily exploit multiple parallel language implementations. We demonstrate the use of GpH, a |Par| Monad, and Eden. We do not re-implement all the parallel internals, as we host this functionality in the |ArrowParallel| type class, which abstracts all parallel implementation logic. The backends can easily be swapped, so we are not bound to any specific one.

@@

%\ptcomment{* We evaluate the expressive power of  Arrow parallelism. Showing that 
%  + there is no need to introduce additional types
%  + that it is possible to define algorthmic skeletons (Section 6,7)
%  + ...

%* We demonstrate that Arrow Parallelism can exploit multiple parallel %Haskell implementations, and with low overhead, i.e. less than X%. 
%... (Section 8) }

%We wrap parallel Haskells inside of our |ArrowParallel| typeclass, but
% why do we aim to abstract parallelism this way and what does this
% approach do better than the other parallel Haskells?
% is such a parallelism abstraction of benefit and to what extent does
% it improve existing approaches?
%\begin{itemize}
%	\item \textbf{Arrow DSL benefits}:
%    To implement parallelism, we do not introduce any new types, but only rely on a typeclass that hosts |parEvalN :: [arr a b] -> arr [a] [b]|, which converts a list of Arrows into a new parallel Arrow. Therefore, we do not lose any benefits of using Arrows as parallelism is just encapsulated in yet another Arrow combinator. The resulting Arrow can be used in the same way a potential serial version could be used. This is a big advantage of this approach, especially compared to Monad solutions like the |Par| Monad which require specialized Monad types. We can just \enquote{plug} in parallel parts into sequential Arrow-based programs without having to change anything.
%	\item \textbf{Abstraction}:
%	With the |ArrowParallel| typeclass, we abstract all parallel implementation logic away from the business logic. This means it is possible to write code against the interface of a common typeclass without being bound to any parallel Haskell. So as an example, during development, we can run the program in a simple GHC-compiled variant and afterwards deploy it on a cluster by converting it into an Eden version, by just replacing the current |ArrowParallel| instance.
%\end{itemize}

%\paragraph{Structure}
%The remaining text is structures as follows. Section~\ref{sec:background} briefly introduces known parallel Haskell flavours (Section~\ref{sec:parEvalNIntro}) and gives an overview of Arrows to the reader (Section~\ref{sec:arrows}). Section~\ref{sec:related-work} discusses related work. Section~\ref{sec:parallel-arrows} defines Parallel Arrows and presents a basic interface. Section~\ref{futures} defines Futures for Parallel Arrows, this concept enables better communication. Section~\ref{sec:map-skeletons} presents some basic algorithmic skeletons  in our newly defined dialect: parallel |map| with and without load balancing. More advanced skeletons are showcased in Section~\ref{sec:topology-skeletons} (|pipe|, |ring|, |torus|). Section~\ref{sec:benchmarks} shows the benchmark results. Section~\ref{sec:conclusion} discusses future work and concludes.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

>>>>

We provide an Arrows-based type class and implementations for the three above mentioned parallel Haskells.
Instead of 
introducing a new low-level parallel backend to implement our
Arrows-based interface, we define a shallow-embedded DSL for Arrows. This DSL
is defined as a common interface with varying implementations in
the existing parallel Haskells. Thus, we not only define a parallel programming interface in a
novel manner -- we tame the zoo of parallel Haskells. We provide a
common, very low-penalty programming interface that allows to switch
the parallel backends at will. The induced penalty was in the single-digit percent range, with means over the varying cores configuration typically under 2\% overhead in our measurements (Section~\ref{sec:benchmarks}). Further backends based on HdpH or a Frege implementation (on the Java Virtual Machine) are viable, too.

@@

\begin{itemize}
\item We introduce a parallel evaluation formalism using Arrows. One big advantage of our specific approach is that we do not have to introduce any new types, facilitating composability (Section~\ref{sec:parallel-arrows}).
\item We show that PArrow programs can readily exploit multiple parallel language implementations. We demonstrate the use of GpH, a |Par| Monad, and Eden. We do not re-implement all the parallel internals, as we host this functionality in the |ArrowParallel| type class, which abstracts all parallel implementation logic. The backends can easily be swapped, so we are not bound to any specific one.

@@
